### 2013-02-12 exam5/1.tex
problem 1

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & 18\\ \hline
2 & & 10\\ \hline
Extra & & 5 \\ \hline
{\bf Total} & & 28+5 \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

@@

In order to manage the distribution of ID , the College decided that they
needed to create a new implementation of the \verb|Set| data
structure.  They went with the lowest bidder, which
turned out to be a bunch of BU students.  Now, having paid for this
code, the College needs to test it, and they turn to you.

Unfortunately, the code the BU students produced is so convoluted that
it doesn't bear looking at.  All you have to go on is that the
\verb|Set| data structure obeys the following interface:

\begin{verbatim}
;; A Set implements:
;; - insert : String -> Set
;;   adds given string to this set, if it isn't already there
;;   (otherwise produces this set)
;; - remove : String -> Set
;;   removes given string from this set, if it's already there
;;   (otherwise produces this set)
;; - size : -> Number
;;   count the elements of this set
;; - has? : String -> Boolean
;;   is the given String in this set?

;; empty-set is a Set value containing no elements
\end{verbatim}
\noindent
\verb|Set|s should satisfy the following property, called the
``insert-size'' property, for any element and
set:

\begin{quote}
If we insert any element into any set, then the resulting set is the
same size as the original set \emph{if} the element was already in the
set, and the resulting set is one larger than the original set
if the element was \emph{not} already in the set.
\end{quote}

\begin{enumerate}
\item Codify the ``insert-size'' property as a predicate.  Write a
  test case that should pass (if the sets work right) using the
  predicate you have defined.
\end{enumerate}

\newpage
\noindent
The College wants to be really sure that their sets are correct, and
they've come up with another property that they think is true, called
the ``insert-remove'' property:

\begin{quote}
If we insert element any element $a$ into any set, and then remove any
element $b$ from the resulting set, the result has the same size as if
we do the two operations in the reverse order, i.e., remove $b$, then
insert $a$.
\end{quote}

\noindent
The College is considering adding this to the test suite as well.

@@

\item Is the ``insert-remove'' property true for all sets?  If so,
  give an explanation.  If not, write a counter-example (i.e., a
  failing test case) using your predicate.

@@

\begin{solution}
\begin{alltt}
\rubricpts{1 pt -- contract & purpose}
;; check-insert-size : Set String -> Boolean
;; check the "insert-size" property on s and i.
\rubricpts{4 pts -- code}
(define (check-insert-size s i)
  (= (s . insert i . size)
     (cond [(s . has? i) (add1 (s . size))]
           [else               (s . size)])))

\rubricpts{2 pts -- test case}
(check-expect (check-insert-size empty-set "hi") true)

\rubricpts{1 pt -- contract & purpose}
;; check-remove-insert : Set String -> Boolean
;; check the "remove-insert" property on s, a, and b.
\rubricpts{3 pts -- code}
(define (check-remove-insert s a b)
  (= (s . insert a . remove b . size)
     (s . remove b . insert a . size))

\rubricpts{1 pt -- successful test}
(check-expect (check-remove-insert empty-set "yes" "no") true)

;; Property does not hold.
;; Counter-example (failing test):
\rubricpts{3 pts -- failing test}
(check-expect (check-remove-insert empty-set "yes" "yes") true)
\end{alltt}
\end{solution}

>>>>

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & 19\\ \hline
2 & & 10\\ \hline
Extra & & 5 \\ \hline
{\bf Total} & & 28+5 \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{19}

@@

In order to manage the distribution of ID cards, the University needs
a data structure for managing sets of numbers.  Van Horn started on
the implementation, resulting in what you see below.

\begin{verbatim}
;; A Set is one of:
;; - (new empty%)
;; - (new singelton% Number)
;; - (new union% Set Set)
;; Interpretation: a set that contains all the *unique* numbers in the data
;; and implements
;; - sum : -> Number
;;   Produce the sum of all the unique elements of this set
;; - sum/acc : Set -> Number
;;   Produce the sum of all the unique elements of this set that are
;;   *not* in the given set
;; - contains? : Number -> Boolean
;;   Does this set contain the given number?
;; - union : Set -> Set
;;   Combine this set with the given set

(define-class empty%
  (define (sum) 0)
  (define (sum/acc s) 0))

(define-class singleton%
  (fields n)
  (define (sum) n)
  (define (sum/acc s) (cond [(s . contains? (this . n)) 0]
                            [else (this . n)])))

(define-class union%
  (fields l r)
  (define (sum) (this . sum/acc (new empty%)))
  (define (sum/acc s) (+ (this . l . sum/acc s)
                         (this . r . sum/acc (s . union (this . l))))))
\end{verbatim}

\newpage

Unfortunately, as you can see, Van Horn fell asleep before finishing,
and the deadline is today.  Help him out by implementing the
\verb|union| and \verb|contains?| methods for all \verb|Set|s.

\ifrubric
\begin{alltt}
\rubricpts{1 pt -- union method}
\rubricpts{3 pts -- contains method}
;; in empty%
(define (union s) s)
(define (contains n) #f)
;; also ok:
(define (union s) (new union% this s)
;; in singleton%
(define (union s) (new union% this s)
(define (contains n) (= n (this . n)))
;; in union%
(define (union s) (new union% this s)
(define (contains n) (or (this . l . contains n)
                         (this . r . contains n)))
(define e (new empty%))
\rubricpts{2 pts -- tests}
(check-expect (e . contains 5) #f)
(check-expect ((new singleton% 5) . contains 5) #t)
(check-expect ((new singleton% 6) . contains 5) #f)
(check-expect ((new singleton% 6) . union  e . contains 5) #f)
(check-expect (e . union (new singleton% 5) . contains 5) #t)
(check-expect ((new singleton% 5) . union (new singleton% 5) . contains 5) #t)
\end{alltt}
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage

\verb|Set|s should satisfy the following property, called the
``union-contains'' property, for any number $n$ and
any two sets $s_1$ and $s_2$:

\begin{quote}
Given two sets $s_1$ and $s_2$, if $n$ is contained in the union of $s_1$ and $s_2$, then $n$ is either
contained in $s_1$ or it is contained in $s_2$.
\end{quote}

\begin{enumerate}
\item Codify the ``union-contains'' property as a predicate.  Write a
  test case that should pass (if your implementations work right) using the
  predicate you have defined.
\end{enumerate}

\ifrubric
\begin{alltt}
\rubricpts{1 pt -- contract & purpose}
;; union-contains : Set Set Number -> Boolean
\rubricpts{3 pts -- code}
(define (union-contains s1 s2 n)
  (check-expect (s1 . union s2 . contains n)
                (or (s1 . contains n) (s2 . contains n))))
\rubricpts{1 pt -- successful test}
(union-contains e (new singleton% 17) 17)
\end{alltt}
\else
\newpage
[Here is some more space for the previous problem.]
\fi

\newpage
\noindent
The University doesn't really trust Van Horn, and they want to make
sure that \verb|Set|s satisfy another property, called the
``union-sum'' property:

\begin{quote}
Given two sets $s_1$ and $s_2$, computing the sum of $s_1$ and the sum
of $s_2$ and then adding the two results is equal to the sum of the
union of  $s_1$ and $s_2$.
\end{quote}

\noindent
The University is considering adding this to the test suite as well.

@@

\item Is the ``union-sum'' property true for all sets?  If so,
  give an explanation.  If not, write a counter-example (i.e., a
  failing test case) using your predicate.

@@

\begin{solution}
\begin{alltt}
\rubricpts{1 pt -- contract & purpose}
;; union-sum : Set Set -> Boolean
\rubricpts{3 pts -- code}
(define (union-contains s1 s2 n)
  (check-expect (s1 . union s2 . sum)
                (+ (s1 . sum) (s2 . sum))))w

\rubricpts{1 pt -- successful test}
;; successful test
(union-sum (new singleton% 17) e)
\rubricpts{3 pts -- failing test}
;; It's false, counterexample:
(union-sum (new singleton% 17) (new singleton% 17))
\end{alltt}
\end{solution}
